---
title: "Covid vaccine data"
author: "Harinath Reddy"
date: "2022-10-26"
output: html_document
self_contained: false
---

```{r}
options(repos = "https://cran.rstudio.com/")

```


```{r}
install.packages("Rtools")
install.packages("readxl")
library("readxl")
```


```{r}


setwd("C:\\Users\\Harinath\\Documents\\GitHub\\Statistical_graphs_and_exploration_using_R")

pfizer <- read_excel("C:\\Users\\Harinath\\Documents\\GitHub\\Statistical_graphs_and_exploration_using_R\\pfizer.xlsx")

Moderna <- read_excel("C:\\Users\\Harinath\\Documents\\GitHub\\Statistical_graphs_and_exploration_using_R\\Moderna.xlsx")


janssen <- read_excel("C:\\Users\\Harinath\\Documents\\GitHub\\Statistical_graphs_and_exploration_using_R\\janssen.xlsx")
```


```{r}

library(tidyverse)
library(dplyr)
```



```{r}
library(devtools)
devtools::install_github('FIRST-Data-Lab/IDDA', force = TRUE)

```

# creating variables from Datasets
```{r}
State <- pfizer$Jurisdiction
Date <- pfizer$Week.of.Allocations
Pfizer.1st.Dose.Allocations <- pfizer$X1st.Dose.Allocations 
Pfizer.2nd.Dose.Allocations <- pfizer$X2nd.Dose.Allocations 
Moderna.1st.Dose.Allocations <- Moderna$X1st.Dose.Allocations
Moderna.2nd.Dose.Allocations <- Moderna$X2nd.Dose.Allocations 
Janssen.1st.Dose.Allocations <- janssen$X1st.Dose.Allocations 
All.Dose.Allocations <- 0.5 *(Pfizer.2nd.Dose.Allocations+Pfizer.1st.Dose.Allocations+Moderna.1st.Dose.Allocations + Moderna.2nd.Dose.Allocations)+Janssen.1st.Dose.Allocations
Cum.Allocation <-cumsum(All.Dose.Allocations)

```


#1.Data Exploration.

# find out the maximum length column, inorder to create a dataframe with equal number of records. As we know different vaccine datasets has different number of records.

```{r}
max_length <- max(c(length(Janssen.1st.Dose.Allocations),length(Moderna.1st.Dose.Allocations),length(Moderna.2nd.Dose.Allocations),length(Pfizer.1st.Dose.Allocations),length(Pfizer.2nd.Dose.Allocations)))
max_length
```


#creating a dummy vaccine_dataframe
```{r}
df_vaccine <- data.frame(Janssen.1st.Dose.Allocations = c(Janssen.1st.Dose.Allocations,
                                                    rep(NA,max_length-length(Janssen.1st.Dose.Allocations))),
                         Moderna.1st.Dose.Allocations = c(Moderna.1st.Dose.Allocations,                                                        rep(NA,max_length-length(Moderna.1st.Dose.Allocations))),
                         Moderna.2nd.Dose.Allocations = c(Moderna.2nd.Dose.Allocations,
                                                    rep(NA,max_length-length(Moderna.2nd.Dose.Allocations))),
                         Pfizer.1st.Dose.Allocations = c(Pfizer.1st.Dose.Allocations,rep(NA, max_length-length(Pfizer.1st.Dose.Allocations))),
                         
                         Pfizer.2nd.Dose.Allocations =c(Pfizer.2nd.Dose.Allocations,
                                                        rep(NA, max_length-length(Pfizer.2nd.Dose.Allocations))),
                         State =c(State, 
                                  rep(NA, max_length-length(State))),
                         Date=c(Date, 
                                  rep(NA, max_length-length(Date))),
                         All.Dose.Allocations = c(All.Dose.Allocations,
                                  rep(NA, max_length-length(All.Dose.Allocations))),
                         Cum.Allocation = c(Cum.Allocation,
                                  rep(NA, ,max_length-length(Cum.Allocation))))
```



```{r}

str(df_vaccine)

```



```{r}
population_data <- IDDA::pop.state
population_data
```


```{r}
population_data  <- population_data%>% rename_at('State', ~'name')



```



# removing the space from state column
```{r}
df_vaccine <- df_vaccine%>%
  mutate(name = sapply(State, gsub, pattern = " ", 
replacement = ""))

```




#creating a left_join using Dplyr
```{r}

df<- left_join(df_vaccine,population_data, by="name")

```

```{r}
str(df)
```


#creating Doseperpop Variable using Mutate

```{r}
library(dplyr)
df <- mutate(df, Doseperpop=Cum.Allocation/df$population)

```
#Making sure to drop all the NA values


```{r}


df <- df[!(df$State=="American Samoa" | df$State=="Palau" | df$State == "Guam" | df$State == "Mariana Islands" | df$State == "Marshall Islands" | df$State == "NewYorkCity" | df$State == "PuertoRico" |  df$State == "U.S.VirginIslands" | df$State == "Philadelphia"),] 


```




```{r}
str(df)
```




#2.Visualization


```{r}
library(plotly)

plot1 <- plot_ly() %>%
  
# Add Cook County’s time series using mode: lines+markers
  add_trace(data = df %>%
              filter(State == 'Virginia'),x = ~Date , y = ~Janssen.1st.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Janssen Dose') %>%
  
   add_trace(data = df %>%
               
               filter(State == 'Virginia'),x = ~Date , y = ~Pfizer.2nd.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Pfizer 2nd dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Pfizer.1st.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Pfizer 1st dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Moderna.2nd.Dose.Allocations, type = 'scatter', mode ='lines+markers',showlegend = TRUE, name = 'Moderna 2nd dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Moderna.1st.Dose.Allocations, type = 'scatter', mode ='lines+markers',showlegend = TRUE, name = 'Moderna 1st dose')

library(htmlwidgets)
saveWidget(plot1, "Figure1.html", selfcontained = F)

```
#Figure 1

```{r}
plot1

```



# (b)
```{r}
plot2 <- plot_ly() %>%

add_trace(data = df %>%
filter(State == "Virginia"),
x = ~Date, y = ~Doseperpop, type = 'scatter', mode = 'lines+markers',
showlegend = TRUE, name = 'Virginia dosage allocation over-time' )

saveWidget(plot2, "Figure2.html", selfcontained = F)
plot2
```
#Figure 2



#(c) 
```{r}
library(plotly)

plot3 <- plot_ly() %>%

add_trace(data = df %>%
group_by(State),                        
x = ~Date, y = ~Doseperpop, type = 'scatter', mode = 'lines+markers',
showlegend = TRUE, name = 'doseperpopulation')
  
saveWidget(plot3, "Figure3.html", selfcontained = F)
plot3
```
#Figure3



#map(d)
```{r}
install.packages("sp")
install.packages("sf")
install.packages("leaflet")
install.packages("geojsonio")
devtools::install_github("rstudio/leaflet")



```
1


```{r}
library(geojsonio); library(leaflet); library(dplyr); library(sp); library(sf)

```



```{r}
library(geojsonio)
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")
class(states)
```



## Merging a spatial and non spatial dataframe using sp::merge 

```{r}
states <- sp::merge(states, df, by="name", duplicateGeoms = T)

```




```{r}

m <- leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))


```



```{r}
bins <- c(0, 25, 50, 75, 100, 125, 150)
pal <- colorBin("YlOrRd", domain = states$Doseperpop, bins = bins)

```


```{r}
labels_cases <- sprintf("<strong>%s</strong><br/>Population: %g M<br>
  Dosage per population: %g<br>
  All dose allocation: %g",
  states$name, round(states$population / (1e6), 2),
  states$Doseperpop, states$All.Dose.Allocations) %>%
 lapply(htmltools::HTML)
labels_cases[[1]]
```



```{r}
m <- m %>% addPolygons(
  fillColor = ~pal(states$Doseperpop),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels_cases,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto"))



m
```


```{r}

m <- m %>% addLegend(pal = pal, values = states$Doseperpop, 
            opacity = 0.7, title = "Dosage per population",
            position = "bottomright")


```
#A high Dosageperpop value indicates that particular state performed poorly in allocating the vaccine doses for it's communities.


```{r}

m

```

```{r}
library(htmltools)
library(htmlwidgets)
saveWidget(m, file="Figure4.html", selfcontained = T)
```


Figure 4

